---
title: "10: Exercise - extend the Shiny dashboard"
format: 
  html: 
    embed-resources: true
editor: visual
---

## Your turn: make the app yours

The lesson walked you through a working Shiny dashboard for a US tech portfolio (AAPL / MSFT / GOOG / AMZN / NVDA). For the exercise, you're going to take that app and extend it in two ways:

1.  **Swap in a new portfolio** of your own choosing - any five tickers, from any sector(s), any weights you like. Pick something you actually care about. An all-ETF portfolio, a sector rotation, a "pandemic winners" basket, a dividend sleeve - your call.
2.  **Add a fifth tab** to the app that adds something new. A few suggestions (pick one, or roll your own):
    -   **Rolling volatility:** plot a 60-day rolling annualized volatility of the portfolio.
    -   **Return histogram:** daily return distribution with VaR and CVaR lines overlaid (borrow from Lesson 8).
    -   **Correlation heatmap:** pairwise correlations of the constituent returns.
    -   **Monte Carlo:** a small Monte Carlo projection of one-year portfolio value based on historical mean/sd.

### What to turn in

-   Your `app.R` as a single file.
-   A short (half a page) write-up: which portfolio you picked, which fifth tab you built, and one interesting thing you noticed when you ran the app (did your portfolio smoke the equal-weight benchmark? Did the forecast look ridiculous? Were the drawdowns scarier than you expected?).

### Starter skeleton

Below is a skeleton that loads libraries, keeps the helper functions from the lesson, and leaves placeholders for you to fill in. Copy this into a new `app.R`, fill in the blanks, and run it.

### Load libraries

```{r}
#| label: load-libraries
#| eval: false

library(shiny)
library(shinyWidgets)
library(tidyverse)
library(tidyquant)
library(plotly)
library(scales)
library(DT)
library(tidymodels)
library(modeltime)
library(timetk)
library(PerformanceAnalytics)
```

### Helpers (same as lesson)

```{r}
#| label: helpers
#| eval: false

perf_summary <- function(r, rf_annual = 0.04) {
    # ... same as lesson 10 ...
}

portfolio_returns <- function(ret_long, weights) {
    # ... same as lesson 10 ...
}
```

### Your portfolio

```{r}
#| label: your-portfolio
#| eval: false

# Pick your 5 tickers and weights here
my_tickers <- c("___", "___", "___", "___", "___")
my_weights <- c(0.2, 0.2, 0.2, 0.2, 0.2)  # must sum to 1
```

### UI

```{r}
#| label: ui-skeleton
#| eval: false

ui <- fluidPage(
    titlePanel("My Portfolio Dashboard"),

    sidebarLayout(
        sidebarPanel(
            # ticker selection, weight inputs, date range
            # (copy from lesson, tweak defaults)
        ),
        mainPanel(
            tabsetPanel(
                tabPanel("Cumulative returns",  plotlyOutput("cum_plot")),
                tabPanel("Drawdown",            plotlyOutput("dd_plot")),
                tabPanel("Performance",         DTOutput("perf_tbl")),
                tabPanel("Forecast",            plotlyOutput("fcst_plot")),
                tabPanel("YOUR NEW TAB",        ___)      # <-- your addition
            )
        )
    )
)
```

### Server

```{r}
#| label: server-skeleton
#| eval: false

server <- function(input, output, session) {

    # reactive price/return pull (copy from lesson)

    # outputs for the first four tabs (copy from lesson)

    # ---- YOUR NEW TAB ----
    output$___ <- ___({
        # your code here
    })
}

shinyApp(ui, server)
```

### Hints

-   If you go with the **rolling volatility** tab, `slider::slide_dbl()` or `zoo::rollapply()` are both fine. Multiply by `sqrt(252)` to annualize.
-   For the **correlation heatmap**, `corrr::correlate() |> stretch()` gives you a long tibble that's easy to `ggplot()` with `geom_tile()`.
-   For **Monte Carlo**, recycle the code from Lesson 8, just drive it from the portfolio's mean and sd of daily returns.
-   For the **return histogram with VaR**, compute VaR and CVaR once at the top of the reactive and add `geom_vline()` for the VaR and a shaded band for CVaR.

Good luck. Once you have it running, show somebody. It's fun to watch non-finance people play with a dashboard they can actually click on.
