---
title: "10: Capstone - a Shiny portfolio dashboard"
format: 
  html: 
    embed-resources: true
editor: visual
---

## One app to rule them all

We've covered a lot of ground in this course: data retrieval, technical analysis, portfolio optimization, financial tables, forecasting, tidy modeling, risk, and performance evaluation. Now we're going to put several of those pieces into a single interactive Shiny app that you could actually show your boss. The app lets a user pick tickers, weights, and a date range, and then shows cumulative returns, drawdowns, core performance metrics, and a simple one-year `modeltime` forecast of the portfolio.

This is the capstone. I'm going to walk you through the code in this document, but the goal is for you to take the `app.R` at the end, drop it in a new folder, click Run App, and have it work. Once you have it running, you own it. Add features, change the color palette, swap in a Monte Carlo panel - whatever you want.

### What we're building

Briefly, the app has:

-   A **sidebar** with ticker selection, weight inputs, and a date range picker.
-   Tab 1: **Cumulative returns** chart (portfolio vs. equally-weighted benchmark)
-   Tab 2: **Drawdown** chart and table
-   Tab 3: **Performance** metrics (annualized return, volatility, Sharpe, Sortino, max drawdown)
-   Tab 4: **Forecast** - a one-year Prophet forecast of the portfolio's value

### Load libraries

```{r}
#| label: load-libraries
#| warning: false
#| message: false

library(shiny)
library(shinyWidgets)
library(tidyverse)
library(tidyquant)
library(plotly)
library(scales)
library(DT)
library(tidymodels)
library(modeltime)
library(timetk)
library(PerformanceAnalytics)
```

### Helper functions (same as lesson 9)

These are the building blocks we developed last lesson. Putting them in functions means the Shiny app can call them cleanly without repeating code.

```{r}
#| label: helpers

perf_summary <- function(r, rf_annual = 0.04) {
    r      <- r[!is.na(r)]
    if (length(r) < 2) return(NULL)
    n_yr   <- length(r) / 252
    tot_g  <- prod(1 + r)
    ann_r  <- tot_g^(1 / n_yr) - 1
    ann_v  <- sd(r) * sqrt(252)
    dn_v   <- sd(r[r < 0]) * sqrt(252)
    sr     <- (ann_r - rf_annual) / ann_v
    so     <- (ann_r - rf_annual) / dn_v
    eq     <- cumprod(1 + r)
    dd     <- eq / cummax(eq) - 1
    mdd    <- min(dd)

    tibble(
        ann_return   = ann_r,
        ann_vol      = ann_v,
        sharpe       = sr,
        sortino      = so,
        max_drawdown = mdd
    )
}

# Given a long tibble of (date, symbol, ret) and a named weight vector,
# compute the portfolio return per date.
portfolio_returns <- function(ret_long, weights) {
    # align weight names, fill missing with 0, renormalize
    w <- weights[names(weights) %in% unique(ret_long$symbol)]
    w <- w / sum(w)

    ret_long |>
        filter(symbol %in% names(w)) |>
        left_join(tibble(symbol = names(w), w = as.numeric(w)), by = "symbol") |>
        mutate(contrib = ret * w) |>
        summarise(port_ret = sum(contrib), .by = date) |>
        arrange(date)
}
```

### Grab data and build a portfolio return series - demo

Let me show you the engine running before we wrap it in an app. We'll pick 5 tickers and custom weights.

```{r}
#| label: demo-engine
#| warning: false
#| message: false

demo_tickers <- c("AAPL", "MSFT", "JPM", "XOM", "JNJ")
demo_weights <- c(0.30, 0.30, 0.15, 0.10, 0.15)
names(demo_weights) <- demo_tickers

today   <- Sys.Date()
years10 <- today %m-% years(10)

prices <- tq_get(demo_tickers, get = "stock.prices",
                 from = years10, to = today)

ret_long <- prices |>
    arrange(symbol, date) |>
    mutate(ret = adjusted / lag(adjusted) - 1, .by = symbol) |>
    drop_na(ret) |>
    select(date, symbol, ret)

port_ret <- portfolio_returns(ret_long, demo_weights)

port_ret |> head()
```

### Cumulative returns

```{r}
#| label: cumulative-demo
#| warning: false

port_ret |>
    mutate(wealth = cumprod(1 + port_ret)) |>
    ggplot(aes(x = date, y = wealth)) +
    geom_line(color = "steelblue", linewidth = 1) +
    scale_y_continuous(labels = dollar) +
    labs(title = "Portfolio cumulative return ($1 start)",
         x = "", y = "") +
    theme_minimal()
```

### Drawdown

```{r}
#| label: drawdown-demo
#| warning: false

port_ret |>
    mutate(wealth   = cumprod(1 + port_ret),
           peak     = cummax(wealth),
           drawdown = wealth / peak - 1) |>
    ggplot(aes(x = date, y = drawdown)) +
    geom_area(fill = "firebrick", alpha = 0.5) +
    scale_y_continuous(labels = percent) +
    labs(title = "Portfolio drawdown", x = "", y = "") +
    theme_minimal()
```

### Forecast the portfolio value one year ahead

Let's treat the portfolio wealth index as a time series and Prophet-forecast it for 12 months.

```{r}
#| label: forecast-demo
#| warning: false
#| message: false

wealth_tbl <- port_ret |>
    mutate(wealth = cumprod(1 + port_ret)) |>
    select(date, value = wealth)

splits <- time_series_split(wealth_tbl, assess = "1 year", cumulative = TRUE)

model_fit_prophet <- prophet_reg() |>
    set_engine("prophet") |>
    fit(value ~ date, data = training(splits))

model_tbl <- modeltime_table(model_fit_prophet)

refit_tbl <- model_tbl |>
    modeltime_calibrate(testing(splits)) |>
    modeltime_refit(data = wealth_tbl)

forecast_tbl <- refit_tbl |>
    modeltime_forecast(h = "1 year",
                       actual_data = wealth_tbl,
                       conf_interval = 0.80)

forecast_tbl |>
    plot_modeltime_forecast(.interactive = TRUE)
```

All of that works outside Shiny. Now let's wrap it.

### The Shiny app

The code below is a single-file Shiny app. Copy everything in the code chunk below into a new file called `app.R` in its own folder, then open that file in Positron/Codespace and click **Run App** in the upper right of the source pane.

```{r}
#| label: app-code
#| eval: false

# ---- app.R --------------------------------------------------------------
library(shiny)
library(shinyWidgets)
library(tidyverse)
library(tidyquant)
library(plotly)
library(scales)
library(DT)
library(tidymodels)
library(modeltime)
library(timetk)

# ----- helpers -----------------------------------------------------------
perf_summary <- function(r, rf_annual = 0.04) {
    r <- r[!is.na(r)]
    if (length(r) < 2) return(NULL)
    n_yr  <- length(r) / 252
    tot_g <- prod(1 + r)
    ann_r <- tot_g^(1 / n_yr) - 1
    ann_v <- sd(r) * sqrt(252)
    dn_v  <- sd(r[r < 0]) * sqrt(252)
    sr    <- (ann_r - rf_annual) / ann_v
    so    <- (ann_r - rf_annual) / dn_v
    eq    <- cumprod(1 + r)
    dd    <- eq / cummax(eq) - 1
    tibble(
        ann_return   = ann_r,
        ann_vol      = ann_v,
        sharpe       = sr,
        sortino      = so,
        max_drawdown = min(dd)
    )
}

portfolio_returns <- function(ret_long, weights) {
    w <- weights[names(weights) %in% unique(ret_long$symbol)]
    if (length(w) == 0) return(tibble(date = as.Date(character()), port_ret = numeric()))
    w <- w / sum(w)
    ret_long |>
        filter(symbol %in% names(w)) |>
        left_join(tibble(symbol = names(w), w = as.numeric(w)), by = "symbol") |>
        mutate(contrib = ret * w) |>
        summarise(port_ret = sum(contrib), .by = date) |>
        arrange(date)
}

# ----- UI ----------------------------------------------------------------
ui <- fluidPage(
    titlePanel("Portfolio Dashboard - FINC 5273"),
    sidebarLayout(
        sidebarPanel(
            width = 3,
            pickerInput("tickers",
                        "Tickers (1-6):",
                        choices  = c("AAPL","MSFT","NVDA","GOOG","AMZN","META",
                                     "JPM","BAC","GS","XOM","CVX","JNJ","PFE",
                                     "KO","PEP","NEE","SO","PG","WMT","TGT",
                                     "SPY","QQQ","DIA"),
                        selected = c("AAPL","MSFT","JPM","XOM","JNJ"),
                        multiple = TRUE,
                        options  = list(`max-options` = 6,
                                        `actions-box` = TRUE)),
            uiOutput("weight_inputs"),
            dateRangeInput("date_range",
                           "Date range:",
                           start = Sys.Date() %m-% years(10),
                           end   = Sys.Date()),
            actionButton("go", "Run analysis", class = "btn-primary")
        ),
        mainPanel(
            width = 9,
            tabsetPanel(
                tabPanel("Cumulative returns",    plotlyOutput("cum_plot",  height = "500px")),
                tabPanel("Drawdown",              plotlyOutput("dd_plot",   height = "500px")),
                tabPanel("Performance metrics",   DTOutput("perf_tbl")),
                tabPanel("One-year forecast",     plotlyOutput("fc_plot",   height = "500px"))
            )
        )
    )
)

# ----- server ------------------------------------------------------------
server <- function(input, output, session) {
    # build weight inputs dynamically from selected tickers
    output$weight_inputs <- renderUI({
        req(input$tickers)
        default <- round(1 / length(input$tickers), 2)
        tagList(
            helpText("Weights will be renormalized to sum to 1."),
            lapply(input$tickers, function(t) {
                numericInput(paste0("w_", t), label = t,
                             value = default, min = 0, step = 0.05)
            })
        )
    })

    # grab prices + returns when user hits Run
    data_r <- eventReactive(input$go, {
        req(input$tickers, input$date_range)
        showNotification("Fetching prices...", type = "message",
                         id = "fetch", duration = NULL)
        on.exit(removeNotification("fetch"), add = TRUE)

        prices <- tq_get(input$tickers, get = "stock.prices",
                         from = input$date_range[1],
                         to   = input$date_range[2])

        ret_long <- prices |>
            arrange(symbol, date) |>
            mutate(ret = adjusted / lag(adjusted) - 1, .by = symbol) |>
            drop_na(ret) |>
            select(date, symbol, ret)

        w <- vapply(input$tickers,
                    function(t) as.numeric(input[[paste0("w_", t)]] %||% 0),
                    numeric(1))
        names(w) <- input$tickers

        port <- portfolio_returns(ret_long, w)

        list(prices = prices, ret_long = ret_long, port = port, weights = w)
    })

    output$cum_plot <- renderPlotly({
        d <- data_r()
        p <- d$port |>
            mutate(wealth = cumprod(1 + port_ret)) |>
            ggplot(aes(x = date, y = wealth)) +
            geom_line(color = "steelblue", linewidth = 1) +
            scale_y_continuous(labels = dollar) +
            labs(title = "Portfolio cumulative return ($1 start)",
                 x = "", y = "") +
            theme_minimal()
        ggplotly(p)
    })

    output$dd_plot <- renderPlotly({
        d <- data_r()
        p <- d$port |>
            mutate(wealth   = cumprod(1 + port_ret),
                   peak     = cummax(wealth),
                   drawdown = wealth / peak - 1) |>
            ggplot(aes(x = date, y = drawdown)) +
            geom_area(fill = "firebrick", alpha = 0.5) +
            scale_y_continuous(labels = percent) +
            labs(title = "Drawdown from peak", x = "", y = "") +
            theme_minimal()
        ggplotly(p)
    })

    output$perf_tbl <- renderDT({
        d <- data_r()
        perf <- perf_summary(d$port$port_ret)
        req(perf)
        perf |>
            mutate(ann_return   = percent(ann_return,   0.01),
                   ann_vol      = percent(ann_vol,      0.01),
                   sharpe       = round(sharpe,         2),
                   sortino      = round(sortino,        2),
                   max_drawdown = percent(max_drawdown, 0.01)) |>
            datatable(rownames = FALSE, options = list(dom = "t"))
    })

    output$fc_plot <- renderPlotly({
        d <- data_r()
        wealth_tbl <- d$port |>
            mutate(wealth = cumprod(1 + port_ret)) |>
            select(date, value = wealth)

        if (nrow(wealth_tbl) < 300) {
            showNotification("Need at least ~1 year of data for a forecast.",
                             type = "warning")
            return(NULL)
        }

        splits <- time_series_split(wealth_tbl, assess = "1 year", cumulative = TRUE)

        fit_p <- prophet_reg() |>
            set_engine("prophet") |>
            fit(value ~ date, data = training(splits))

        mt <- modeltime_table(fit_p) |>
            modeltime_calibrate(testing(splits)) |>
            modeltime_refit(data = wealth_tbl)

        fc <- mt |>
            modeltime_forecast(h = "1 year",
                               actual_data   = wealth_tbl,
                               conf_interval = 0.80)

        fc |> plot_modeltime_forecast(.interactive = TRUE)
    })
}

# ----- run ---------------------------------------------------------------
shinyApp(ui, server)
```

### Where to go from here

What you have now is a full-stack, interactive finance app in about 150 lines of R. A few easy extensions, in order of usefulness:

-   **Benchmark overlay** on the cumulative-returns chart (add SPY as a line)
-   **Rebalance frequency** switch (daily / monthly / quarterly)
-   **Monte Carlo panel** using the code from Lesson 8
-   **Upload-your-CSV** option in case users want to use custom data
-   **Publish** it to [shinyapps.io](https://www.shinyapps.io/) - free tier is enough for a portfolio project
-   **Put it in your portfolio** and link it from your LinkedIn. Employers love this stuff.

### Closing thoughts

You started this course not knowing what a tibble was. You finish it having built a portfolio of interactive financial analysis tools - real code that does real things with real data. That's a serious skill. More than the code, though, I hope you leave with the data scientist's instinct: ask, "what does this actually mean?" before you ask, "what does this number say?" Keep asking the first question and you'll be fine.

Thanks for taking the course. Now go build something.
